---
title: "Obtain Fishbase/Sealifebase Data"
output: html_document
date: "2024-12-25"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load in Packages

```{r}
# For Packages for fishbase
remotes::install_github("cboettig/duckdbfs", force = TRUE)
remotes::install_github("ropensci/rfishbase")
library(duckdbfs)
library(rfishbase)

# Writing to .parquet file (faster than .csv)
library(arrow)
```
**IMPORTANT / SCRIPT PURPOSE:** The purpose of this script is to obtain fishbase and sealifebase data from both the country() and faoareas() functons, containing their species names. The original country() and faoareas() only contains the species codes, and thus must be joined to the species codes to obtain species names, which is necessary for interfacing it with the ARTIS data.

# Obtain fishbase / sealifebase data

```{r  create fb/slb data}
if (!file.exists("../data/fb_slb_data/")) {
  # Obtain FB / SLB country() data
  fb_presences_country <- country(server = "fishbase")
  slb_presences_country <- country(server = "sealifebase")
  
  fb_species_codes <- load_taxa(server = "fishbase") %>%
    rename(ScientificName = "Species") %>%
    mutate(Species = sapply(str_split(ScientificName, " "), `[`, 2)) %>%
    relocate(Species, .before = Genus)

  slb_species_codes <- load_taxa(server = "sealifebase") %>%
    rename(ScientificName = "Species") %>%
    mutate(Species = sapply(str_split(ScientificName, " "), `[`, 2)) %>%
    relocate(Species, .before = Genus)
  
  # Join FB / SLB country() to species codes
  fb_presences_country <- fb_presences_country %>%
    left_join(fb_species_codes, by = "SpecCode") %>%
    rename(SubFamily = "Subfamily") %>%
    mutate(Phylum = NA_character_, Kingdom = NA_character_) %>%
    relocate(c(ScientificName, Species, Genus, SubFamily, Family, Order, Class, SuperClass, Phylum, Kingdom), .after = SpecCode)
  
  slb_presences_country <- slb_presences_country %>%
    left_join(slb_species_codes, by = "SpecCode") %>%
    rename(SubFamily = "Subfamily") %>%
    mutate(SuperClass = NA_character_) %>%
    relocate(c(ScientificName, Species, Genus, SubFamily, Family, Order, Class, SuperClass, Phylum, Kingdom), .after = SpecCode)
  
  # Obtain FB / SLB faoareas() data
  fb_presences_faoareas <- faoareas(species_list = fb_species_codes$Species, server = "fishbase")
  slb_presences_faoareas <- faoareas(species_list = fb_species_codes$Species, server = "sealifebase") 
  
  # Create new directory for FB/SLB data
  dir.create("../data/fb_slb_data")
  
  # Write files to .parquet
  write_parquet(fb_presences_country, "../data/fb_slb_data/fb_presences_country.parquet")
  write_parquet(slb_presences_country, "../data/fb_slb_data/slb_presences_country.parquet")
  write_parquet(fb_presences_faoareas, "../data/fb_slb_data/fb_presences_faoareas.parquet")
  write_parquet(slb_presences_faoareas, "../data/fb_slb_data/slb_presences_faoareas.parquet")
  write_parquet(slb_species_codes, "../data/fb_slb_data/fb_species_codes")
  write_parquet(slb_species_codes, "../data/fb_slb_data/slb_species_codes.parquet")
}
```

> Last date obtained: 1/13/2025